---
title: "Optimizing abyss with optimx"
author: "Hamid Younesy"
date: "October 16, 2016"
output:
  html_document:
    keep_md: yes
---

## Introduction

**["optimx"](https://cran.r-project.org/web/packages/optimx/index.html)** is a 
general-purpose optimization wrapper function that calls other R tools for
optimization.

### Usage
```{r, eval=FALSE}
optimx(par, # a vector of initial values for the parameters for which optimal values are to be found. 
       fn,  # A function to be minimized, with first argument the vector of parameters
       method=c("Nelder-Mead","BFGS"),  # list of the methods to be used (can be more than one)
       lower=-Inf, upper=Inf, # Bounds on the variables
       itnmax=NULL, # maximum number of iterations
       control=list(),
       ...)
```

Possible method codes are: 'Nelder-Mead', 'BFGS', 'CG', 'L-BFGS-B',
'nlm', 'nlminb', 'spg', 'ucminf', 'newuoa', 'bobyqa', 'nmkb', 'hjkb', 'Rcgmin', or 'Rvmmin'.

We have is a wrapper function testOptimx() around optimx to run optix for a given function and
plot the results.

```{r, echo=FALSE, eval=TRUE, results='hide'}
library(optimx)
library(ggplot2)

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

testOptimx <- function(func, init_param, findMaximum=FALSE) {
    df_all <- NULL
    ans_all <- NULL
    for (mtd in c('Nelder-Mead', 'BFGS', 'CG', 'L-BFGS-B'
                  , 'nlm', 'nlminb', 'spg', 'ucminf', 'newuoa' 
                  , 'bobyqa', 'nmkb', 'hjkb', 'Rcgmin', 'Rvmmin'
                  )) {
        results <<- NULL
        ncall <<- 1
        ans <- optimx(
            fn = function(x) {
                r <- func(x)
                results <<- rbind(results, c(ncall, r, x))
                ncall <<- ncall + 1
                r
                }
            , method = mtd,
            #, lower=-20, upper=20
            , par = init_param
            , itnmax = 30
            , control=list(maximize=findMaximum, save.failures=TRUE, trace=0)
            )
        df_all <- rbind(df_all, cbind(data.frame(results), mtd))
        ans_all <- rbind(ans_all, ans)
    }

    colnames(df_all) <- c("n", "metric", paste("p",1:length(init_param), sep=""), "method")
    ans_all["method"] <- rownames(ans_all)

    return (list(df_all = df_all, ans_all = ans_all))
}

plotOptimx <- function(optimxResults, 
                       metricLim = NULL, p1Lim = NULL,
                       metricLabel="metric", p1Label="p1") {
    df_all <- optimxResults$df_all
    ans_all <- optimxResults$ans_all
    metric_stats = boxplot.stats(df_all$metric)$stats
    p1_stats = boxplot.stats(df_all$p1)$stats

    if (is.null(metricLim)) {
        metricLim[1] <- metric_stats[1]
        metricLim[2] <- metric_stats[5]
    }
    if (is.null(p1Lim)) {
        p1Lim[1] <- p1_stats[1]
        p1Lim[2] <- p1_stats[5]
    }

    p1 <- ggplot(df_all, aes(color=method)) +
        theme_bw() +
        geom_line(aes(x=n, y=p1)) +
        ylim(p1Lim[1], p1Lim[2]) +
        xlab("call index") +
        ylab(p1Label) +
        theme(legend.key.size = unit(0.5, "lines"), legend.text = element_text(size = rel(0.5)))

    #print(p)

    
    p2 <- ggplot(df_all, aes(color=method)) +
        theme_bw() +
        geom_line(aes(x=n, y=metric)) +
        ylim(metricLim[1], metricLim[2]) +
        xlab("call index") +
        ylab(metricLabel) +
        theme(legend.key.size = unit(0.5, "lines"), legend.text = element_text(size = rel(0.5)))
    #print(p)

    p3 <- ggplot(ans_all, aes(x=method, y=value)) +
        theme_bw() + 
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        geom_segment(aes(xend = method, yend = 0), size = 10) +
        ylim(metricLim[1], metricLim[2]) +
        xlab("method") +
        ylab(paste("best", metricLabel))
    #print(p)
    
    #multiplot(p1, p2, p3, cols=1)
    print(p1)
    print(p2)
    print(p3)
}
```

## Testing with simple functions.

First, let's test some simple functions of the form f(x, y) = f1(x) + f2(y).
These will all be minimization and the optimum metric is 0.
To mimick discrete parameters we have currently opted for rounding things so that fir an int x, the optimizer queries between (x-. 5,x+.5] get passed as x to the function. since k parameter in abyss is still always int. 
Interestingly convergence did not happen for most algorithms in optimx's package if we were floor'ing parameters

### Two continious parameters: f(x1, x2) = abs(x1 - 5) + abs(x2 - 4)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
optimxResult <- testOptimx(
    func = function(x){abs(x[1]-5) + abs(x[2]-4)}
    , init_param = c(0, 0)
    )

plotOptimx(optimxResult
    , metricLim=c(0, 50), p1Lim=c(-20, 20)
    , metricLabel="f(x1,x2)", p1Label="x1"
    )
```

### One continious and one discrete parameter: f(x1, x2) = abs(x1 - 5) + abs(round(x2)-4)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
optimxResult <- testOptimx(
    func = function(x){abs(x[1]-5) + abs(round(x[2])-4)}
    , init_param = c(0, 0)
    )

plotOptimx(optimxResult
    , metricLim=c(0, 50), p1Lim=c(-20, 20)
    , metricLabel="f(x1,x2)", p1Label="x1"
    )
```

### Two discrete parameters: f(x1, x2) = abs(round(x1)-5) + abs(round(x2)-4)

```{r, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
optimxResult <- testOptimx(
    func = function(x){abs(round(x[1])-5) + abs(round(x[2])-4)}
    , init_param = c(0, 0)
    )

plotOptimx(optimxResult
    , metricLim=c(0, 50), p1Lim=c(-20, 20)
    , metricLabel="f(x1,x2)", p1Label="x1"
    )
```

## Testing with abyss
Data: hsapien chr3, 500k reads

Varying paramter: k

Fixed parameters: s = 200, l = 25

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide'}
setwd("/Users/hyounesy/SFU/Hackseq2016/2016_project_2")
source(paste(getwd(), "scripts/R_optimizers/common/AbyssWrapper.R", sep="/"))
optimxResult <- testOptimx(
    func = function(x){runAbyss500k(round(x[1]))}
    , init_param = c(20, 0)
    , findMaximum = TRUE 
    )

plotOptimx(optimxResult
    , metricLim=c(0, 30000), p1Lim=c(0, 50)
    , metricLabel="N50", p1Label="k"
    )
```



