---
title: "Optimizing abyss with optimx"
author: "Hamid Younesy"
date: "October 16, 2016"
output: pdf_document
---

## Introduction

**["optimx"](https://cran.r-project.org/web/packages/optimx/index.html)** is a 
general-purpose optimization wrapper function that calls other R tools for
optimization.

### Usage
```{r, eval=FALSE}
optimx(par, # a vector of initial values for the parameters for which optimal values are to be found. 
       fn,  # A function to be minimized, with first argument the vector of parameters
       method=c("Nelder-Mead","BFGS"),  # list of the methods to be used (can be more than one)
       lower=-Inf, upper=Inf, # Bounds on the variables
       itnmax=NULL, # maximum number of iterations
       control=list(),
       ...)
```

Possible method codes are: 'Nelder-Mead', 'BFGS', 'CG', 'L-BFGS-B',
'nlm', 'nlminb', 'spg', 'ucminf', 'newuoa', 'bobyqa', 'nmkb', 'hjkb', 'Rcgmin', or 'Rvmmin'.

We have is a wrapper function testOptimx() around optimx to run optix for a given function and
plot the results.

```{r, echo=FALSE, eval=TRUE}
library(optimx)
library(ggplot2)

testOptimx <- function(func, init_param, findMaximum=FALSE, 
                       metricLim = NULL, p1Lim = NULL ) {
    df_all <<- NULL
    ans_all <<- NULL
    for (mtd in c('Nelder-Mead' , 'BFGS', 'CG', 'L-BFGS-B'
                  , 'nlm', 'nlminb', 'spg', 'ucminf', 'newuoa' 
                  , 'bobyqa', 'nmkb', 'hjkb', 'Rcgmin', 'Rvmmin'
                  )) {
        results <<- NULL
        ncall <<- 1
        ans <- optimx(
            fn = function(x) {
                r <- func(x)
                results <<- rbind(results, c(ncall, r, x))
                ncall <<- ncall + 1
                r
                }
            , method = mtd,
            #, lower=-20, upper=20
            , par = init_param
            , itnmax = 30
            , control=list(maximize=findMaximum, save.failures=TRUE, trace=0)
            )
        df_all <<- rbind(df_all, cbind(data.frame(results), mtd))
        ans_all <<- rbind(ans_all, ans)
    }

    colnames(df_all) <<- c("n", "metric", paste("p",1:length(init_param), sep=""), "method")
    
    metric_stats = boxplot.stats(df_all$metric)$stats
    p1_stats = boxplot.stats(df_all$p1)$stats

    
    
    if (is.null(metricLim)) {
        metricLim[1] <- metric_stats[1]
        metricLim[2] <- metric_stats[5]
    }
    if (is.null(p1Lim)) {
        p1Lim[1] <- p1_stats[1]
        p1Lim[2] <- p1_stats[5]
    }

    p <- ggplot(df_all, aes(color=method)) +
        theme_bw() +
        geom_line(aes(x=n, y=metric)) +
        coord_cartesian(ylim = c(metricLim[1], metricLim[2]))
    print(p)

    p <- ggplot(df_all, aes(color=method)) +
        theme_bw() +
        geom_line(aes(x=n, y=p1)) +
        coord_cartesian(ylim = c(p1Lim[1], p1Lim[2]))
    print(p)

    barplot(ans_all$value
            , names=rownames(ans_all)
            , ylim=metricLim
            , ylab = "best metric"
            , las=2)
}
```

## Testing with simple functions.

First, let's test some simple functions of the form f(x, y) = f1(x) + f2(y).
To mimick discrete parameters we will use round().

### Two continious parameters
```{r, echo=TRUE, warning=FALSE, message=FALSE}
testOptimx(
    func = function(x){abs(x[1]-5) + abs(x[2]-4)}
    , init_param = c(0, 0)
    , metricLim=c(0, 50), p1Lim=c(-20, 20)
    )
```

### one discrete parameter
```{r, echo=TRUE, warning=FALSE, message=FALSE}
testOptimx(
    func = function(x){abs(round(x[1])-4)}
    , init_param = c(0, 0)
    , metricLim=c(0, 50), p1Lim=c(-20, 20)
    )
```

### One continious and one discrete parameter
```{r, echo=TRUE, warning=FALSE, message=FALSE}
testOptimx(
    func = function(x){abs(x[1]-5) + abs(round(x[2])-4)}
    , init_param = c(0, 0)
    , metricLim=c(0, 50), p1Lim=c(-20, 20)
    )
```

### Two discrete parameters
```{r, echo=TRUE, warning=FALSE, message=FALSE}
testOptimx(
    func = function(x){abs(round(x[1])-5) + abs(round(x[2])-4)}
    , init_param = c(0, 0)
    , metricLim=c(0, 50), p1Lim=c(-20, 20)
    )
```

## Testing with abyss 200k data
```{r, echo=TRUE, warning=FALSE, message=FALSE}
source(paste(getwd(), "scripts/R_optimizers/common/AbyssWrapper.R", sep="/"))
testOptimx(
    func = function(x){runAbyss200k(round(x[1]))}
    , init_param = c(20, 0)
    , findMaximum = TRUE 
    , metricLim=c(0, 10000), p1Lim=c(0, 50)
    )
```



