---
title: "Hackseq.Rplots"
author: "Veera"
date: "17 October 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 3-Dimensional plots 

```{r}
#load the file 
dfm = read.table("iterations.tsv", sep = "\t", header = TRUE)

#load the libraries
library(scatterplot3d)

#create a data frame only with the 3 columns to be plotted
bar.dfm <- dfm[,c("k","l","N50")]

#define the plot 
s3d <- scatterplot3d(bar.dfm, type = "h", color = grey(nrow(bar.dfm):1/40), 
                     angle = 55, scale.y = 1,pch = 16)

#add linear model
my.lm <- lm(bar.dfm$N50 ~ bar.dfm$k + bar.dfm$l)

#plot the linear model
s3d$plane3d(my.lm)
```
library(ggplot2)
library(reshape2)

dfm.mat = acast(dfm, k ~ l, value.var = "N50")

dfm.mat[ is.na(dfm.mat)] = 0

dfm.melt = melt(dfm.mat)
dfm.melt$Var1 = as.factor(dfm.melt$Var1)
dfm.melt$Var2 = as.factor(dfm.melt$Var2)
dfm.melt.m = ddply(dfm.melt, .(Var2), transform, rescale = rescale(value))

ggplot(dfm.melt.m, aes(Var2,Var1)) + geom_tile(aes(fill = value), colour = "white")


dfm.data.matrix = data.matrix(dfm.mat)
dfm_heatmap = heatmap(dfm.data.matrix, Rowv=NA, Colv=NA, col = cm.colors(256), margins=c(5,10))
